---
title: "Example script"
output: 
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float:
      toc_collapsed: true
    theme: united
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{example-script}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

## Load data

```{r load}
# Setup options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Clear environment
rm(list = ls())

# Load relevant libraries
library(modelLong)
library(dplyr)    # Data manipulation
library(ggplot2)  # Data visualization

# Load example longitudinal data (must be of long format)
ortho <- read.csv("../data/ortho.csv")[-1]
ortho <- ortho %>%
  dplyr::mutate(Subject = as.factor(Subject)) %>%  # Make subject ID a factor
  dplyr::arrange(Subject)  # Arrange measures by subject
ortho_labelled <- ortho

# Rename variables for (pretty printing) 
ortho_labelled <- ortho %>% dplyr::rename("Distance" = distance, "Age" = age)

# Preview data
ortho_labelled[sample(nrow(ortho_labelled), 5),]
```

## Data Summary

Print a summary table of the orthodontist data.

```{r summary, warning=FALSE}
# Data summary, stratified by Sex
summary_res <- modelLong::summary(
    data = ortho_labelled[-1], # Ignore Subject ID
    group_var = "Sex",
    type = list(Age ~ "continuous"))

summary_res$table1 
```

## Data Visualization

Plot repeated measures of adolescent tooth growth over time.

```{r plots}
# Create plot of repeated measures
plot_res <- modelLong::plot(
  data = ortho, 
  outcome = "distance", 
  time = "age", 
  id = "Subject",
  group_var = "Sex")

# Plot repeated measures
plot_res$data_plot +
  # Adjust labels and x-axis
  labs(title = "Repeated measures of tooth growth") +
  xlab("Age (years)") + ylab("Distance (mm)") +
  scale_x_continuous(breaks = seq(8,18,2))

# Plot autocorrelation
plot_res$acf_plot +
  # Adjust labels
  labs(title = "Autocorrelation of tooth growth") + xlab("Age (years)")
```

## Data Modeling

# Marginal effects with GEE

Estimate the mean tooth growth associated with each sex population. This is suitable for a linear regression model fit with generalized estimating equations to account for the fact that our data is repeated measures over time; ignoring clusters and time may bias our estimates.

```{r gee}
# Fit a GEE model of adolescent tooth growth from age and sex
id <- as.factor(ortho$Subject)
mod_gee <- modelLong::fit_gee(
  data = ortho, 
  formula = formula(distance ~ age + Sex), 
  id = id,
  family = gaussian(), 
  corstr = "independence")

# Fitted coefficients
coef(mod_gee$fit)

# Full model
mod_gee$fit
```

# Conditional effects with GLMM

Estimate the mean tooth growth associated with each individual. This is well modeled by a  linear regression model fit with mixed-effects to account each subject in the data.

```{r glmm}
# Fit a GLMM model of adolescent tooth growth from age and sex
mod_glmm <- modelLong::fit_glmm(
  data = ortho,
  formula = formula(distance ~ age + Sex + (1 | Subject)),
  family = gaussian)

# Fitted coefficients
coef(mod_glmm$fit)

# Full model
mod_glmm$fit
```

